### 2022-03-15 paper/localpsf.tex
restructuring paper

The results from Section \ref{sec:intromoments} allow us to compute $V$, $\mu$, and $\Sigma$ by applying $\Aop^T$ to a small number of polynomial functions. Since a probability distribution concentrates within a small number of standard deviations of its mean, knowing $\mu(x)$ and $\Sigma(x)$ permits us to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$ (Section \ref{eq:mean_and_covariance_estimation}). This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}). We compute one batch of $\impulseresponse_{x_i}$'s by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a collection of points $x_i$ scattered throughout $\Omega$. The batch of points $x_i$ are chosen via a greedy ellipsoid packing algorithm so that the support of $\impulseresponse_{x_i}$ and the support of $\impulseresponse_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$ (Section \ref{sec:sample_point_selection}). Because these supports do not overlap, we can post-process the response of $\Aop$ to the Dirac comb to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i$ in the batch---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section \ref{sec:get_impulse_response}). The process is repeated to get more batches, until a desired number of impulse responses is reached.

Once the scattered impulse responses $\impulseresponse_{x_i}$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation (Section \ref{sec:approximate_kernel_entries}). The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section \ref{eq:local_mean_displacement_invariance}.
%Specifically, we make the assumption that
%\begin{equation}
%\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}(y - \mu(x) + \mu(x_i))
%\end{equation}
%if $x_i$ is close to $x$. This is shown in Figure \ref{fig:mean_displacement_invariance}. Motivation for this assumption is discussed in Section \ref{eq:local_mean_displacement_invariance}. 
We approximate $\Phi(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\frac{V(x)}{V(x_i)}\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for a collection of sample points $x_i$ near $x$. The weights are determined by radial basis function interpolation.

@@

When $\Aop$ is positive semi-definite, $\mathbf{A}_H$ may be indefinite due to errors in the approximation. In this case, we modify the H-matrix representation of $\mathbf{A}_H$ to make it positive semi-definite, by applying a specially chosen rational matrix function to $\mathbf{A}_H$. This is described in Appendix \ref{sec:make_spd}.

Schematically the overall H-matrix approximation proceeds as follows:
\begin{equation*}
\Aop \longrightarrow V, \mu, \Sigma \longrightarrow \{x_i\}_{i=1}^r \longrightarrow \underbrace{\{\impulseresponse_{x_i}\}_{i=1}^r}_{\text{as }\eta_b\text{'s}} \longrightarrow \boldsymbol{\Phi}_H \longrightarrow \mathbf{A}_H.
\end{equation*}
The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm \ref{alg:construct_Atilde}.

@@

\subsubsection{Cost to evaluate one approximate kernel entry}

>>>>

The results from Section \ref{sec:intromoments} allow us to compute $V$, $\mu$, and $\Sigma$ by applying $\Aop^T$ to a small number of polynomial functions. Since a probability distribution concentrates within a small number of standard deviations of its mean, knowing $\mu(x)$ and $\Sigma(x)$ permits us to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$ (Section \ref{eq:mean_and_covariance_estimation}). This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}). We compute the impulse response batch $\eta_b$ by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a batch, $S_b$, of points $x_i$ scattered throughout $\Omega$ (Section \ref{sec:get_impulse_response}). The batch of points $S_b$ is chosen via a greedy ellipsoid packing algorithm so that, for $x_i,x_j \in S_b$, the support of $\impulseresponse_{x_i}$ and the support of $\impulseresponse_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$ (Section \ref{sec:sample_point_selection}). Because these supports do not overlap, we can post-process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section \ref{sec:get_impulse_response}). The process is repeated to get more batches, until a desired number of impulse responses is reached.

Once the impulse response batches $\eta_b$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation (Section \ref{sec:approximate_kernel_entries}). The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section \ref{eq:local_mean_displacement_invariance}.
%Specifically, we make the assumption that
%\begin{equation}
%\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}(y - \mu(x) + \mu(x_i))
%\end{equation}
%if $x_i$ is close to $x$. This is shown in Figure \ref{fig:mean_displacement_invariance}. Motivation for this assumption is discussed in Section \ref{eq:local_mean_displacement_invariance}. 
We approximate $\Phi(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\frac{V(x)}{V(x_i)}\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for a collection of sample points $x_i$ near $x$. The weights are determined by radial basis function interpolation.

@@

When $\Aop$ is positive semi-definite, $\mathbf{A}_H$ may be indefinite due to errors in the approximation. In this case, one may (optionally) modify the H-matrix representation of $\mathbf{A}_H$ to make it positive semi-definite, by applying a specially chosen rational matrix function to $\mathbf{A}_H$. This is described in Appendix \ref{sec:make_spd}.

Schematically the overall H-matrix approximation proceeds as follows:
\begin{equation*}
\Aop \longrightarrow V, \mu, \Sigma \longrightarrow \underbrace{\{x_i\}_{i=1}^r}_{\text{as }S_b\text{'s}} \longrightarrow \underbrace{\{\impulseresponse_{x_i}\}_{i=1}^r}_{\text{as }\eta_b\text{'s}} \longrightarrow \boldsymbol{\Phi}_H \longrightarrow \mathbf{A}_H.
\end{equation*}
The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm \ref{alg:construct_Atilde}. The cost to construct $\mathbf{A}_H$ is discussed in Section \ref{sec:overall_cost}.

@@
